\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-27 metatheory-condensed.tex
Comment.

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
%
Formation rules for substitutions have the form
$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
valid in the context~$\Dl$ yields terms valid in~$\Gm$.
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.

>>>>

As is standard in MLTT presentations, we use explicit substitutions~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are represented by de Bruijn indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{(\lam{\var{1}})}$.
%
Formation rules for substitutions have\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) } the form 
$\goodSub{\Gm}{\gm}{\Dl}$, which says that applying~$\gm$ to terms
valid in the context~$\Dl$ yields terms valid in~$\Gm$.
%
The two main forms of substitutions are weakening and extension:
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/proj} states that if $t$ is a dependent pair
of type $\TySigma{A}{B}$, then $\snd{t}$ has type $\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$.
